---
title: "Assignment1"
author: "Harshini Balam"
date: "2024-02-05"
output: pdf_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, comment = NA)
```
1. The Assignment consists of the data regarding the IMDB rating, which inculdes both quantitative and qualitative variables. Dataset Source: https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows
```{r}
library(readr)
library(knitr)
library(tinytex)
library(latexpdf)
movies <- read_csv("C:/Users/Harshini/Downloads/imdb_top_1000.csv")
```

```{r}
View(movies)
```

2. Descriptive Statistics for the variables “Runtime" and “IMDB_Rating”
```{r}
summary(movies[,c('Runtime' , "IMDB_Rating")])
```
3. Transforming the “IMDB_Rating” variable into Square Root
```{r}
square_IMDB_Rating <- (movies$IMDB_Rating)^(1/2)
square_IMDB_Rating
```
4. Plotting the variables “Meta Score” and “IMDB_Rating” against each other
```{r}
plot(movies$Meta_score, movies$IMDB_Rating, main= "Meta Score vs. IMDB_Rating", xlab='MetaScore', ylab='IMDB_Rating')
```
5.Scatter Plot for the variables “Meta Score” and “IMDB_Rating”
```{r}
library(ggplot2)
ggplot(data=movies, aes(x=Meta_score, y=IMDB_Rating)) + geom_point(color = "red", size = 2) +
  labs(title = "Scatter Plot of metascore vs. IMDB_Rating")
```

